---
title: "kongerækken, homework"
author: "simone kallehauge"
date: "2022-11-28"
output: html_document
---
#indlæs libraries 
```{r}
library(tidyverse)
library(here)
```


#indlæs data 
```{r}
kongeraekken <- read_csv2("../DM03/data/kongerække_homework_csv2.csv")
```



#data visualisering
my how-to: 
  i've first selected the name and the duration of the reign from my Excel spreadsheet,
  but in order to to make sure that RStudio order them by reign and not first-name-letter, 
  i make a new column in the spreedsheet, that states what order in the reign they are. 
    here: name_count
  
  next i mutate my data, because R sees the "duration of reign" as a character, not as a number, 
  which then makes my visulization wrong. So in order to make sure it order after the numeric system, 
  i use the function: as.numeric
  
  i then in ggplot reorder my data by name and name_count, (see select)
  
  in order to make the monarchs name readable, i use the function: theme(axis.text.x = element_text(angle=90)) <-- this makes it stand verticle 
  
  lastly i organize the y-axis by using the "scale_y_continuous(breaks = seq), which makes me able to use lesser numbers 
    (0 = the smallest possible number) 
    (65 = the highest possible number) 
    (5= the breaks: 0,5,10,15,20 etc.)
  
```{r}
kongeraekken %>% 
  select(name, Duration_Of_Reign,name_count) %>%
  mutate(Duration_Of_Reign= as.numeric(Duration_Of_Reign)) %>% 
  ggplot(aes(x= reorder(name, name_count), y=Duration_Of_Reign))+
  geom_col()+
  theme(axis.text.x = element_text(angle=90))+
  labs(title = "kongerækken", 
       y = "Reign Duration", 
       x = "Monarch")+
  scale_y_continuous(breaks = seq(0, 65, 5))

```



