---
title: "casualties caused by police exam"
author: "Simone Broni Kallehauge"
date: "2023-01-09"
output:
  pdf_document: default
  html_document: default
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Read libraries needed}
library(htmltools)
library(tidyverse)
library(dplyr)
library(leaflet)
library(plotly)
library(ggplot2)
```

#Create the data link to data: <https://mappingpoliceviolence.org/>

```{r Read in the data}
police_killing_2022 <- read_csv("Mapping-Police-Violence-DIVIDED DATA.csv")
```

The data i use here, is the cleaned data from OpenRefine, where the date
is seperated into day, month, year.

##Create the Cluster Map of victims in 2022

I've found the following code in Brent Thornes video:
<https://www.youtube.com/watch?v=dBk8gGX1MNk> (09/12/2022)

```{r Ensuring the coordinates are numeric}
police_killing_2022$longitude <- as.numeric(police_killing_2022$longitude)
police_killing_2022$latitude <- as.numeric(police_killing_2022$latitude)
```

Even though the coordinates in the data frame all are written
numerically, R can sometimes misinterpret the data, and view it as a
character. If the data we want to use to make a map is viewed as
characters, the code won't work, and the map won't show. So in order to
ensure that the data is infact in numeric, we can use the code written
above.

The reason for using "\$": - to the left of \$: the dataframe we want to
use. the right of the \$: the column from the dataset It works almost
like an intermediater: direct R to where it should get the information
from (but not like "\<-")


```{r making the coordinates their own value}
long <- police_killing_2022[[43]]
lat <- police_killing_2022[[42]]
```

The reason for making the longitude and latitude their own value, is
that (as you will see in the code-chuck with leaflet [the map]) it makes
the coding easier, thus i don't have to plot in the coordinates
separately.

    Description:
      long/lat = the name of the new value 
      <- police_killing = the place that i create my value from 
      [[42/43]] = the column in which the new value gets it's value/data from

```{r Filtering the year used for the Cluster Map}
police_killing_2022 %>% 
  filter(year == 2022)
```

Usually, when using the "filter" function, you have to surround the data
in quotation marks, but because the data I want to filter is numeric,
using quotation marks isn't necessary A way to find out if the data you
want to use is either numeric or character, you can simply upen the data
frame (will open a new tap) and then hover the mouse over the name of
the column.

```{r making a clustered map of the killings - without racial division}
leaflet() %>% 
  addTiles() %>%
  addMarkers(lng = long, 
             lat = lat,
             popup = paste(police_killing_2022$race,'<br>',  
                           police_killing_2022$name, '<br>', 
                           police_killing_2022$age, '<br>', 
                           police_killing_2022$gender, '<br>', 
                           police_killing_2022$city),
             clusterOptions = markerClusterOptions())
```

If you click on one of the clusters, it will spread out to smaller
clusters, until you reach the singular marker, which here is a blue
popup. On the singular popup, there is some information about the
victim: the race, the name, the age, and the city it happend.
Unfortunately, all the information listed above, isn't always
available - due to unknown reasons (e.g. an unknown male from Salem)
Another flaw with this visualization, is that it is not possible to
separate the clusters in race, so in order to figure out how many that
actually gets killed, there has to be another form of visualization

    Tutorial used for this visulization:
      <https://rstudio.github.io/leaflet/markers.html> (used 06/01/2023)

Description of the code used [Map]: - leaflet is the package. - AddTiles
() makes the map. - the lng and lat in "addMarkers" is respectively
longitude and latitude. Instead of plotting in the coordinated one by
one, I've then made a value (*as you can see in the upper grey-code,
which is also described above the grey code*), which leads back to the
longitude and latitude in my dataset. I've found the code clusterOptions
here: <https://rstudio.github.io/leaflet/markers.html>

It appears that for some reason, Rstudio won't load my map correctly, so it shows all the casualities from 2013-2022. But in my final report can you see the map as it should be seen

```{r count the number of killings by race, 2022}
police_killing_2022 %>% 
  count(race, sort = TRUE)
```

The reasons for me to count how many victims there were in 2022, is that
it is a good way to get a good overview of the numbers; in this case how
many casualties there were in 2022, but also beacuse it is useful in
another form of visulization:

###Number of killings, divided by race - but not Population percentage
For this visualization, I've used some of the code we worked with during
the week 48 lesson: Webscraping (What I've used from this, is the
filtering of race, seen in line three)

```{r Visualising the killings in the different states in 2022}
illustration_1 <- police_killing_2022 %>% 
  select(state, year, race) %>% 
  filter(race %in% c("Black", "White", "Hispanic", "Native American", "Unknown race")) %>% 
  filter(year == "2022") %>% 
  count(state, race) %>% 
  rename(Victims ="n") %>% 
  ggplot(aes(x=state,
             y= Victims,
             fill=race))+
  geom_col()+
   theme(axis.text = element_text(angle = 90))

ggplotly(illustration_1)
```

##Killings in the US done by police divided by race, population in state

```{r read in the dataset about us states population}
library(readxl)
us_states_code<- read_excel("NST-EST2022-POP. ALTERED DATA.xlsx")
```

undgå videnskabelig notation i % angivelse:

```{r}
options(scipen= 999)
```

```{r Counting the victims by race in each state}
police_killing_2022 %>% 
  select(state, year, race) %>% 
  filter(race %in% c("Black", "White", "Hispanic", "Native American", "Unknown race")) %>% 
  filter(year == "2022") %>% 
  count(state, race)
```

```{r rename the column name to state}
us_states_code <- us_states_code %>% 
  rename(state = state_code)
```

The reason for this, is that in order for left-join (see below) to work,
the data that need to be joined has how have the same data-name. So in
this case, in the police_killing DF and the us_states DF the column
containing the state abbreviations is named the same

```{r calculating the percentages of victims by race in each state}
police_killing_2022 %>% 
  select(state, year, race) %>% 
  filter(race %in% c("Black", "White", "Hispanic", "Native American", "Unknown race")) %>% 
  filter(year == "2022") %>% 
  count(state, race) %>% 
  
  left_join(us_states_code, by="state") %>% 
  mutate(percentage_killed = n / pop) 
```

A disclaimer: The total number of population in each state is repeated,
so the number of pop by each race does not mean that it is the number of
e.g. black population in the state. However the percentage of victims
(race) in each state - is correct.

### Visualization of victims in each state, including percentage

```{r}
col_2 <- police_killing_2022 %>% 
  select(state, year, race) %>% 
  filter(race %in% c("Black", "White", "Hispanic", "Native American", "Unknown race")) %>% 
  filter(year == "2022") %>% 
  count(state, race) %>% 
  left_join(us_states_code, by="state") %>% 
  mutate(percentage_killed = n / pop) %>% 
  rename(Victims ="n") %>% 
  ggplot(aes(x=state,
             y= percentage_killed,
             fill=race))+
  geom_col()+
  theme(axis.text = element_text(angle = 90))

ggplotly(col_2)

```

```{r counting the number of victims - by race }
police_killing_2022 %>%  
select(state, year, race) %>% 
  count(race, sort=TRUE)
```

```{r calculating which state has the highest number of black victims}
police_killing_2022 %>%  
select(state, year, race) %>% 
  filter(year == "2022") %>% 
  filter(race == "Black") %>% 
  count(state, sort=TRUE)
```

```{r calculating which state has the highest number of white victims}
police_killing_2022 %>%  
select(state, year, race) %>% 
  filter(year == "2022") %>% 
  filter(race == "White") %>% 
  count(state, sort=TRUE)
```

```{r}
sessionInfo()
```
